---
title: "Tidyverse Train the Trainer"
author: "Day 2"
date: "1/16/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Notes

- [Make it Stick: Science of Succesful Learning](https://www.amazon.com/Make-Stick-Science-Successful-Learning/dp/0674729013)
- Teaching `readr` later in the day gives a buffer section. It's someting that can be dropped as a "self-contained" unit, or if on time can be explained. Several people in the class discussed that it can be distracting to show people how to import data because they take off from there. 
- Instead, starting off by writing code can be an excellent introduction. copy/paste, run code, then start modifying the code. Gets students "coding" immediately.

## Workshop Outlines

1. Two Day Workshop
  - Day 1
    - Import & Intro
    - Visualize
    - Transform Part 1
    - Transform Part 2 and Output
  - Day 2
    - Review (import %>% transform %>% visualize)
    - Tidy
    - Scripts, workflow
    - Project


## Mental Models

- Tidy data
  - Bento box
  - pick and choose lunch menu
  

## Serparate and Separate Rows

- building on the idea of "tidy" data: one unit of observation per cell
- sometimes cells are overflowing with data, imagine an oveflowing cupboard
- what you want in the end are groups of like things: cups in one place, plates in another
- `separate()` and `separate_row()` are like adding cupboards or adding shelves
- You use `separate()` when you need new cupboard (move plates out of the cup cupboard), `separate_row()` when you need more shelves (unstack cups)
- Both functions have the same form...
- by default the separation operates on anything that isn't a letter or number
- but you can change this as well

## ADEPT

- Analogy
- Diagram
- Example
- Plain english
- Technical definition

## Excercises

- Start building exercises early in the preparation process
- Put together useful data sets
- scaffolding: choose a level of difficulty that matches the learner's progress
  - reading code that's disorganized and comment on reason

## My `tidyr` learning example

`tidyr` is part of the tidyverse

```{r}
library(tidyverse)
```

Run this code chunk below. (Don't look inside, just run it.)

```{r}
starswars <- 
  starwars %>% 
  mutate(
    hair_eyes = paste(hair_color, eye_color, sep = ";"), 
    films = map_chr(films, paste, collapse = ";")
  ) %>% 
  filter(!is.na(hair_color)) %>% 
  select(1, hair_eyes, films) %>% 
  slice(1:5)
```

You now have a table called `starswars`

```{r}
starswars
```

Notice that both `hair_eyes` and `films` are "overflowing".
The `hair_eyes` column contains two separate concetps: hair color and eye color.
The `films` column contains multiple films in a single cell.

```r
<separate>(data, <column>, <?into>, sep = "<separator>")
```

```{r}
separate_rows(starswars, films, sep = ";")
```

```{r}
separate(starswars, hair_eyes, into = c("hair", "eyes"), sep = ";")
```

Change the name of the columns where hair color and eye color were stored to `hair_color` and `eye_color`.

Re-unite hair color and eye color, separated by `/` instead of `,`.

```{r}
starswars %>% 
  separate(hair_eyes, into = c("hair", "eyes"), sep = ";") %>% 
  unite(hair_eyes, hair, eyes, sep = "/")
```